---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

### Data Preprossessing
```{r, echo = F, include = F}
library(data.table)
library(tidyverse)
library(gridExtra)

### Import the dataset and set the order of cols for fast retrieval.

loanStat2017 <- lapply(Sys.glob("data/LoanStats_2017Q*.csv"), fread, na.strings = "", skip = 1)
names(loanStat2017) <- paste0("Q", 1:4)
invisible(lapply(loanStat2017, setcolorder, neworder = order(names(loanStat2017$Q1))))

ls2017 <- rbindlist(loanStat2017)

### 1. Numeric, Categorical and Date Columns-------------------------------------------------


numCols <- c("acc_now_delinq", "acc_open_past_24mths", "all_util", "annual_inc", "annual_inc_joint",
             "avg_cur_bal", "bc_open_to_buy", "bc_util", "chargeoff_within_12_mths", 
             "collection_recovery_fee", "collections_12_mths_ex_med", "delinq_2yrs", "delinq_amnt",
             "dti", "dti_joint", "funded_amnt", "funded_amnt_inv", 
             "il_util", "inq_fi", "inq_last_12m", "inq_last_6mths",
             "last_pymnt_amnt", "loan_amnt", "max_bal_bc", "mo_sin_old_il_acct",
             "mo_sin_old_rev_tl_op", "mo_sin_rcnt_rev_tl_op", "mo_sin_rcnt_tl", "mort_acc", 
             "mths_since_last_delinq", "mths_since_last_major_derog", "mths_since_last_record",
             "mths_since_rcnt_il", "mths_since_recent_bc", "mths_since_recent_bc_dlq",
             "mths_since_recent_inq", "mths_since_recent_revol_delinq", "num_accts_ever_120_pd", 
             "num_actv_bc_tl", "num_actv_rev_tl", "num_bc_sats", "num_bc_tl", 
             "num_il_tl", "num_op_rev_tl", "num_rev_accts", "num_rev_tl_bal_gt_0", 
             "num_sats", "num_tl_120dpd_2m", "num_tl_30dpd", "num_tl_90g_dpd_24m",
             "num_tl_op_past_12m", "open_acc", "open_acc_6m", "open_act_il", 
             "open_il_12m", "open_il_24m", "open_rv_12m", "open_rv_24m", "orig_projected_additional_accrued_interest",
             "out_prncp", "out_prncp_inv", "pct_tl_nvr_dlq", "percent_bc_gt_75", 
             "pub_rec", "pub_rec_bankruptcies", "revol_bal", "revol_bal_joint", 
             "sec_app_chargeoff_within_12_mths", "sec_app_collections_12_mths_ex_med",                     
             "sec_app_inq_last_6mths", "sec_app_mort_acc",                          
             "sec_app_mths_since_last_major_derog", "sec_app_num_rev_accts",                     
             "sec_app_open_acc", "sec_app_open_act_il", "sec_app_revol_util",
             "tax_liens", "tot_coll_amt", "tot_cur_bal", "tot_hi_cred_lim",
             "total_bal_ex_mort", "total_bal_il", "total_bc_limit", "total_cu_tl", "total_il_high_credit_limit",
             "total_pymnt", "total_pymnt_inv", "total_rec_int", "total_rec_late_fee", 
             "total_rec_prncp", "total_rev_hi_lim")
ls2017[, (numCols) := lapply(.SD, as.numeric), .SDcols = numCols]
## Warning "In lapply(.SD, as.numeric) : NAs introduced by coercion": owe to  as.numeric("NA")

# Be espeicially carefully on int_rate, in the form of "4.7%"
ls2017[, ":="(int_rate = as.numeric(gsub("%", "", int_rate)), 
              revol_util = as.numeric(gsub("%", "", revol_util)))]

# str(ls2017[, 90:ncol(ls2017)])
dateCols <- c("debt_settlement_flag_date", "earliest_cr_line", 
              "hardship_end_date", "hardship_start_date", 
              "issue_d", "last_credit_pull_d", "last_pymnt_d", "next_pymnt_d", 
              "payment_plan_start_date", "sec_app_earliest_cr_line",
              "settlement_date")
# ls2017[, dateCols, with = F]
ls2017[, (dateCols) := lapply(.SD, parse_date, format = "%b-%Y"), .SDcols = dateCols]

```

## First Plot 
```{r, warning=F}
## .1
jointPos <- grep("_joint", names(ls2017))
ls2017.joint <- ls2017[, names(ls2017)[c(jointPos, jointPos - 1)], with = F]
setcolorder(ls2017.joint, order(names(ls2017.joint)))
# str(ls2017.joint)

# violin default already scaled: 
## Income and Revolving Balance has a long tail, which is believe to be outliers (false data)
## One possible method is to use log(x+1) transform
ls2017.joint[, paste0(names(ls2017.joint)[1:6], "_log") := log(.SD + 1), .SD = names(ls2017.joint)[1:6]]


p11 <- ggplot(melt(ls2017.joint[,9:10], measure.vars = names(ls2017.joint[,9:10])), 
              aes(x = variable, y = value)) + geom_violin() + 
  ggtitle("Density Plot of Annual Income")


p12 <- ggplot(ls2017.joint, aes(x = dti_log, y = dti_joint_log)) + geom_hex() + 
  ggtitle("Binary Plot of dti")

p13 <- ggplot(melt(na.omit(ls2017.joint[,7:8]), measure.vars = names(ls2017.joint[,7:8])), 
              aes(x = variable, y = value)) + geom_count() + ggtitle("Verification Status Count")

## .2
secAppColnames <- grep("sec_app_", names(ls2017), value = T)
ls2017.secApp <- ls2017[, c(gsub("sec_app_", "", secAppColnames), secAppColnames), with = F]

p14 <- ggplot(melt(na.omit(ls2017.secApp[, c(3,13)]), measure.vars = names(ls2017.secApp)[c(3,13)]), 
              aes(x = value, y = ..density.., col = variable)) + geom_freqpoly() + 
  ggtitle("Frequency Polygon of Earlies Credit Line Date")

grid.arrange(p11, p12, p13, p14, ncol = 2)
```

For the features with info both of borrower, one can see the similarity. 
The coborrower’s  income is higher, which might be the case parents are the co-borrower of the students. The meaning of dti is a ratio of debt payment and income. Co-borrower has a low ratio. They don't have a huge burden. 
The verification status is somehow interesting. Less co-borrowers are verified might due to the verification mechanism of Lending Club. 
Second applicants built a late credit line. They are elder. So they contact credit products later. 



### Second Plot

```{r, echo = F}
ls2017.joint.log <- ls2017.joint[, -(1:6)]


# Create coborrower's class
ls2017.joint.log[, cbClass := is.na(verification_status_joint) + 
                   is.na(annual_inc_joint_log) + 
                   is.na(dti_joint_log) + 
                   is.na(revol_bal_joint_log)]
ls2017.joint.log[, cbType := ifelse(cbClass == 0, "none", ifelse(cbClass == 4, "all", "med"))]
ls2017.joint.log[, ":="(int_rate = ls2017$int_rate, loan_status = ls2017$loan_status, 
                        loan_amnt = ls2017$loan_amnt, grade = ls2017$grade)]

## let the class be balanced to do better visualization 
## cbType: none missing(10%) in "_joint", 1~3 missing(1%), all missing(89%)
a <- table(ls2017.joint.log$cbType)

set.seed(101)
allInd_inv <- sample(which(ls2017.joint.log$cbType == "all"), a[1] - a[3])


p21 <- ggplot(ls2017.joint.log[-allInd_inv], aes(x = annual_inc_log, y = int_rate, col = cbType)) + 
  geom_point() + geom_jitter(height = 2, width = 0, cex = .1) + 
  ggtitle("Interest Rate VS Annual Income (with or without Coborrowers")



p22 <- ggplot(ls2017.joint.log[-allInd_inv], aes(x = verification_status, y = cbType)) + 
  geom_raster(aes(fill = int_rate), hjust = .5, vjust = .5, interpolate = F) + 
  ggtitle("Interest Rate VS Verification Status (with or without Coborrowers")
grid.arrange(p21, p22, ncol = 1)
```

The cbType, referring to co-borrower's type, is defined as follows. All means the applicant has a co-borrower, none means no, and med denotes the applicant's co-borrower information is partially missing. 
If we add one response, e.g. interest rate, one can see a separate pattern against annual income under the cbType. 
As for verification status, the trends differs a lot among these 3 cbTypes.

A casual linear regression is also applied and result in a slight decrease of the prediction error. 

